---
title: "Disruption Testing"
output: 
  html_document:
    self_contained: true
    toc: true
    toc_depth: 3
    toc_float:
      collapsed: true
      smooth_scroll: false
params:
  base_scen: 'C:\FTOT\scenarios\reference_scenarios\rs7_capacity'
---

<!-- Add the following line manually to the rendered HTML document so that IE does not block the javascript elements: -->
<!-- saved from url=(0014)about:internet --> 

```{r setup, include=FALSE, warning=FALSE, message=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
options(warn = -1) # Suppress warnings

use_lib <- ifelse(any(grepl("FTOTnetworkEnv", .libPaths())),
  .libPaths()[grepl("FTOTnetworkEnv", .libPaths())],
  .libPaths()
)

suppressPackageStartupMessages(library(ggplot2, lib.loc = use_lib))
suppressPackageStartupMessages(library(dplyr, lib.loc = use_lib))
suppressPackageStartupMessages(library(tidyr, lib.loc = use_lib))
suppressPackageStartupMessages(library(DT, lib.loc = use_lib))
suppressPackageStartupMessages(library(plotly, lib.loc = use_lib))
suppressPackageStartupMessages(library(knitr, lib.loc = use_lib))

scen_name = strsplit(basename(params$base_scen), '_')[[1]][1]
```

# Plotting the disruption scenarios

Total scenario cost versus disruption steps. The number of edges (roadway segments) used at each step is shown by color of the step.

```{r BC_disrupt}
disrupt_type = 'BC'

disrupt_root = paste(params$base_scen, disrupt_type, 'disrupt', sep = '_')

res = read.csv(file.path(disrupt_root, 'Results.csv'))

# clean up commas
res$total_cost <- as.numeric(sub('\\,', '', res$total_cost))

g1 <- ggplot(res, aes(x = disrupt_step, 
                      y = total_cost)) +
  geom_step(size = 2, color = 'grey80') +
  geom_point(size = 2, aes(color = nedge)) +
  theme_bw() +
  ggtitle(paste('Betweenness centrality disruption scenario:', scen_name))

ggsave(plot = g1, filename = paste0(scen_name, '_BC_Disruption_Cost.jpg'))

ggplotly(g1)

```

## Relationship of betweenness centrality and volume

Using the edges in the optimal solution.

Both panels show the relationship between betweenness centrality and background commodity volume flow for the links used in the optimal solution for this scenario.

The left panel colors in the points by the rank in terms of volume. The top links are colored in.
The right panel colors in the points by the rank in terms of sum of betweenness centrality. The top links (used for the disruption test) are colored in.

```{r bc_vol, fig.width=10}
edges <- read.csv(file.path(params$base_scen, 'Edges_to_Remove.csv'))

e_bc <- edges[order(edges$sum_BC, decreasing = T),] %>%
  mutate(BC_rank = order(sum_BC, decreasing = T)) %>%
  mutate(BC_rank = ifelse(BC_rank <= 50, BC_rank, NA))

e_v <- edges[order(edges$volume, decreasing = T),] %>%
  mutate(V_rank = order(volume, decreasing = T)) %>%
  mutate(V_rank = ifelse(V_rank <= 50, V_rank, NA))

e_j <- full_join(e_bc, e_v)

g2_v <- ggplot(e_j, aes(x = volume, 
                      y = sum_BC)) +
  geom_point(aes(color = V_rank), size = 3) +
  theme_bw() +
  ggtitle('Betweenness centrality sum and Volume of edges \n in optimal scenario solution')

g2_bc <- ggplot(e_j, aes(x = volume, 
                      y = sum_BC)) +
  geom_point(aes(color = BC_rank), size = 3) +
  theme_bw() #+
  #ggtitle('Betweenness centrality sum and Volume of edges in optimal scenario solution')

egg::ggarrange(g2_v, g2_bc, ncol = 2)

# Zoom in on just the links identified by either BC or V

top_edges = e_j %>% filter(!is.na(BC_rank) | !is.na(V_rank))

# top_edges %>% View()

```

## Table of segment links removed

Table of segment links which were removed in the disruption steps.

```{r edge_table}

datatable(e_j  %>% filter(!is.na(BC_rank) | !is.na(V_rank)) %>%
            select(BC_rank,
                   V_rank,
                   edge_id,
                   length,
                   capacity,
                   volume,
                   sum_BC,
                   mode_oid),
          caption = paste("Road segments (edges) disrupted in", scen_name),
          rownames = F,
          filter = 'top',
          options = list(dom = "ftp",
                         pageLength = 10)
          ) %>% formatRound(4:5, digits = 2) 

```
